\documentclass{article}
\usepackage[polutonikogreek,english]{babel}
\usepackage[utf8x]{inputenx}
\newcommand{\greek}[1]{{\selectlanguage{polutonikogreek}#1}}

\title{Wine drinking habits in ancient Greece\\ \textit{\small Essay for the course on Greek and Roman Mythology @ Penn}}
\author{Vitaly Repin}
\date{May 2013}

\begin{document}
\maketitle
\newpage
\large
Ancient Greeks almost never drank pure wine. It was diluted with water (or, sometimes, snow). They strongly believed that only barbarians drank undiluted wine.

Drinking habits acted as clear and visible line between civilization and barbarism for the ancient Greeks.  They even had special verb (\textit{\greek{ἐπισκυθίζω}}) in the language
with the meaning \textit{``pour out drink in Scythian fashion, i.e.with unmixed wine''.}

This is also articulated in Greek mythology.
Let's open Homer's masterpiece --- Odyssey.  In the book IX, Maron, a spared priest of Apollo from Ismarus, gave to Odyssey the irresistible wine, ``drink fit for the gods'', and Homer describes how
this wine was expected to be consumed (Fagles, p.~142): \textit{\textbf{twenty cups of water} he’d stir in one of wine.} This is how civilized Greeks enjoy great wine.
Cyclops are not shown as civilized (Fagles, p.~139):

\begin{quote}
\it
They have no meeting place for council, no laws either,
no, up on the mountain peaks they live in arching caverns ---
each a law to himself, ruling his wives and children,
not a care in the world for any neighbor
\end{quote}

After reading this, we can not be surprised that these barbarians also drink undiluted wine. Odyssey used the knowledge of barbarian drinking habits against
Polyphemus (Fagles, p.~146): \textit{three bowls I brimmed and three he drank to the last drop.} The result of barbarian way of drinking is expected --- ogre becomes ``heavy with wine''
(greek word \textit{\greek{οἰνοβαρείων}} is used by Homer), Odyssey and his people are able to blind the monster.

This example illustrates the idea of functionalism applied to mythology --- myths legitimize the cultural norms of the society which is telling the myth. In case of Odyssey the Greeks
are telling to themselves that it is unsafe to drink the undiluted wine and use Polyphemus as an example. He was drinking ``drink fit for the gods'' in improper way and instead of  enjoyment
became blind.

\end{document}
